---
title: "Hypothesis Testing for a Single Mean"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(ggcheck)
library(tidyverse)
library(broom)
library(infer)

gradethis::gradethis_setup()

# Sample data for exercises
set.seed(123)
coffee_data <- tibble(
  customer_id = 1:50,
  daily_cups = rnorm(50, mean = 2.8, sd = 1.2)
)

# Another dataset for practice
study_hours <- tibble(
  student_id = 1:30,
  hours_per_week = rnorm(30, mean = 18.5, sd = 4.2)
)
```

## Introduction

Welcome to this tutorial on hypothesis testing for a single mean! In this lesson, you'll learn how to:

- Set up null and alternative hypotheses
- Calculate test statistics
- Determine p-values
- Make statistical conclusions
- Use R and the tidyverse to perform these tests

### Learning Objectives

By the end of this tutorial, you will be able to:

1. Formulate appropriate null and alternative hypotheses for a single mean
2. Calculate and interpret a t-test statistic
3. Use R functions to perform one-sample t-tests
4. Make decisions based on p-values and significance levels

## Topic 1: Understanding Hypothesis Testing

Hypothesis testing is a statistical method used to make decisions about population parameters based on sample data.

### The infer Workflow

The infer package provides a consistent grammar for statistical inference that follows this pattern:

```
data %>%
  specify(response = variable) %>%
  hypothesize(null = "point", mu = value) %>%
  calculate(stat = "t")
```

This approach makes the logic of hypothesis testing explicit and easier to understand.

### Example Scenario

A coffee shop claims that their customers drink an average of 3 cups of coffee per day. You collect data from 50 customers to test this claim using the infer workflow.

```{r explore-data, exercise=TRUE}
# Let's look at our coffee data
coffee_data %>%
  glimpse()

# Calculate basic summary statistics
coffee_data %>%
  summarise(
    n = n(),
    mean_cups = mean(daily_cups),
    sd_cups = sd(daily_cups),
    se_cups = sd(daily_cups) / sqrt(n())
  )
```

```{r explore-data-check}
grade_code("Great! You've successfully explored the structure and basic statistics of our coffee data.")
```

## Topic 2: Setting Up Hypotheses

### Step 1: Formulate Your Hypotheses

For our coffee shop example:

$$
H_{0}: \mu = 3 \quad \text{vs} \quad H_{A}: \mu \neq 3
$$

### Practice: State the Hypotheses

Consider a university that claims students study an average of 20 hours per week. You want to test if this claim is accurate.

```{r hypotheses-quiz, echo=FALSE}
question("What should the null hypothesis be?",
  answer("H₀: μ < 20", message = "This would be for a one-sided test where we specifically think the mean is less than 20."),
  answer("H₀: μ > 20", message = "This would be for a one-sided test where we specifically think the mean is greater than 20."),
  answer("H₀: μ = 20", correct = TRUE, message = "Correct! The null hypothesis states that the population mean equals the claimed value."),
  answer("H₀: μ ≠ 20", message = "This looks like an alternative hypothesis, not a null hypothesis."),
  allow_retry = TRUE
)
```

# ```{r alternative-quiz, echo=FALSE}
question("If we want to test whether the true mean is different from 20 (either higher or lower), what should the alternative hypothesis be?",
  answer("H₁: μ = 20", message = "This is actually the null hypothesis."),
  answer("H₁: μ ≠ 20", correct = TRUE, message = "Correct! This is a two-sided alternative hypothesis."),
  answer("H₁: μ < 20", message = "This would be correct for a one-sided test if we specifically suspected the mean was less than 20."),
  answer("H₁: μ > 20", message = "This would be correct for a one-sided test if we specifically suspected the mean was greater than 20."),
  allow_retry = TRUE
)
```

## Topic 3: The infer Grammar of Inference

### Understanding the infer Workflow

The infer package breaks down hypothesis testing into clear, logical steps:

1. **specify()**: Identify the variable of interest
2. **hypothesize()**: State the null hypothesis
3. **calculate()**: Compute the test statistic
4. **get_p_value()**: Calculate the p-value

### Step-by-Step with Coffee Data

Let's work through each step of the infer workflow:

```{r infer-specify, exercise=TRUE}
# Step 1: Specify the response variable
coffee_specified <- coffee_data %>%
  specify(response = daily_cups)

# Look at the structure
coffee_specified
```

```{r infer-specify-check}
grade_code("Great! You've specified the response variable using infer's specify() function.")
```

### Adding the Null Hypothesis

```{r infer-hypothesize, exercise=TRUE}
# Step 2: Add the null hypothesis
coffee_hypothesis <- coffee_data %>%
  specify(response = daily_cups) %>%
  hypothesize(null = "point", mu = 3)

# View the result
coffee_hypothesis
```

```{r infer-hypothesize-check}
grade_code("Perfect! You've added the null hypothesis that the true mean equals 3.")
```

### Calculating the Test Statistic

```{r infer-calculate, exercise=TRUE}
# Step 3: Calculate the t-statistic
coffee_t_stat <- coffee_data %>%
  specify(response = daily_cups) %>%
  hypothesize(null = "point", mu = 3) %>%
  calculate(stat = "t")

# Display the t-statistic
coffee_t_stat
```

```{r infer-calculate-check}
grade_code("Excellent! You've calculated the t-statistic using the infer workflow.")
```

### Practice: Complete infer Workflow

Now try the complete infer workflow for the study hours data:

```{r practice-infer-workflow, exercise=TRUE}
# Complete the infer workflow for study hours
# Test H₀: μ = 20 vs H₁: μ ≠ 20

study_t_stat <- study_hours %>%
  specify(response = ___) %>%
  hypothesize(null = ___, mu = ___) %>%
  calculate(stat = ___)

study_t_stat
```

```{r practice-infer-workflow-solution}
study_t_stat <- study_hours %>%
  specify(response = hours_per_week) %>%
  hypothesize(null = "point", mu = 20) %>%
  calculate(stat = "t")

study_t_stat
```

```{r practice-infer-workflow-check}
grade_code("Perfect! You've mastered the infer workflow for calculating test statistics.")
```

## Topic 4: Getting P-values with infer

### The get_p_value() Function

Once we have our observed test statistic, we can calculate the p-value using `get_p_value()`:

```{r infer-pvalue, exercise=TRUE}
# Calculate p-value for coffee data (two-sided test)
coffee_p_value <- coffee_data %>%
  specify(response = daily_cups) %>%
  hypothesize(null = "point", mu = 3) %>%
  calculate(stat = "t") %>%
  get_p_value(obs_stat = ., direction = "two-sided")

coffee_p_value
```

```{r infer-pvalue-check}
grade_code("Great! You've calculated the p-value using infer's get_p_value() function.")
```

### Complete Test with infer

We can also do the entire test in one pipeline and store the statistic for p-value calculation:

```{r complete-infer-test, exercise=TRUE}
# Complete hypothesis test using infer
observed_t_stat <- coffee_data %>%
  specify(response = daily_cups) %>%
  hypothesize(null = "point", mu = 3) %>%
  calculate(stat = "t")

# Get the p-value
p_value_result <- coffee_data %>%
  specify(response = daily_cups) %>%
  hypothesize(null = "point", mu = 3) %>%
  calculate(stat = "t") %>%
  get_p_value(obs_stat = observed_t_stat, direction = "two-sided")

# Display results
print("Observed t-statistic:")
observed_t_stat

print("P-value:")
p_value_result
```

```{r complete-infer-test-check}
grade_code("Excellent! You've performed a complete hypothesis test using the infer framework.")
```

### Practice: Your Turn with Study Hours

Complete the hypothesis test for study hours using infer:

```{r practice-complete-infer, exercise=TRUE}
# Test if students study significantly different from 20 hours per week
# Complete both the test statistic and p-value calculation

# Step 1: Calculate observed t-statistic
study_observed <- study_hours %>%
  specify(response = ___) %>%
  hypothesize(null = ___, mu = ___) %>%
  calculate(stat = ___)

# Step 2: Calculate p-value
study_p_value <- study_hours %>%
  specify(response = ___) %>%
  hypothesize(null = ___, mu = ___) %>%
  calculate(stat = ___) %>%
  get_p_value(obs_stat = study_observed, direction = ___)

# Display results
print("Observed t-statistic:")
study_observed

print("P-value:")
study_p_value
```

```{r practice-complete-infer-solution}
study_observed <- study_hours %>%
  specify(response = hours_per_week) %>%
  hypothesize(null = "point", mu = 20) %>%
  calculate(stat = "t")

study_p_value <- study_hours %>%
  specify(response = hours_per_week) %>%
  hypothesize(null = "point", mu = 20) %>%
  calculate(stat = "t") %>%
  get_p_value(obs_stat = study_observed, direction = "two-sided")

print("Observed t-statistic:")
study_observed

print("P-value:")
study_p_value
```

```{r practice-complete-infer-check}
grade_code("Outstanding! You've successfully completed a full hypothesis test using the infer workflow.")
```

## Topic 5: Interpreting Results and Making Decisions

### Understanding infer Output

When using infer, you get clean, interpretable output:

- **Test statistic**: The calculated t-value
- **P-value**: A tibble showing the probability
- **Clear workflow**: Each step is explicit and logical

### Comparing infer to Traditional Methods

The infer approach has several advantages:
- **Consistency**: Same grammar for different types of tests
- **Clarity**: Each step is explicit
- **Flexibility**: Easy to switch between parametric and simulation-based methods
- **Tidyverse integration**: Works seamlessly with dplyr and ggplot2

### Decision Rules

- If p-value ≤ α (typically 0.05): **Reject H₀**
- If p-value > α: **Fail to reject H₀**

### Practice: Interpretation

```{r interpretation-quiz}
question("If your t-test gives a p-value of 0.03 and you're using α = 0.05, what should you conclude?",
  answer("Fail to reject the null hypothesis", message = "Not quite. Compare the p-value (0.03) to α (0.05)."),
  answer("Reject the null hypothesis", correct = TRUE, message = "Correct! Since p-value (0.03) < α (0.05), we reject H₀."),
  answer("Accept the null hypothesis", message = "We never 'accept' the null hypothesis, only fail to reject it."),
  answer("The test is inconclusive", message = "The test gives us a clear decision based on the p-value."),
  allow_retry = TRUE
)
```

## Topic 6: Comprehensive Analysis with infer

Let's put it all together with a complete analysis using the infer workflow:

```{r complete-infer-analysis, exercise=TRUE}
# Complete analysis of coffee consumption data using infer

# Step 1: Explore the data (using tidyverse)
coffee_summary <- coffee_data %>%
  summarise(
    n = n(),
    mean_cups = mean(daily_cups),
    sd_cups = sd(daily_cups),
    se_cups = sd(daily_cups) / sqrt(n())
  )

print("Data Summary:")
coffee_summary

# Step 2: Hypothesis test using infer workflow
# H₀: μ = 3 vs H₁: μ ≠ 3

# Calculate test statistic
observed_stat <- coffee_data %>%
  specify(response = daily_cups) %>%
  hypothesize(null = "point", mu = 3) %>%
  calculate(stat = "t")

# Calculate p-value
p_value <- coffee_data %>%
  specify(response = daily_cups) %>%
  hypothesize(null = "point", mu = 3) %>%
  calculate(stat = "t") %>%
  get_p_value(obs_stat = observed_stat, direction = "two-sided")

print("Hypothesis Test Results:")
print(paste("T-statistic:", round(observed_stat$stat, 4)))
print(paste("P-value:", round(p_value$p_value, 4)))

# Step 3: Decision
alpha <- 0.05
decision <- ifelse(p_value$p_value < alpha, 
                   "Reject the null hypothesis", 
                   "Fail to reject the null hypothesis")
print(paste("Decision (α = 0.05):", decision))

# Step 4: Create a visualization
coffee_plot <- coffee_data %>%
  ggplot(aes(x = daily_cups)) +
  geom_histogram(bins = 15, fill = "skyblue", alpha = 0.7, color = "black") +
  geom_vline(xintercept = 3, color = "red", linetype = "dashed", size = 1) +
  geom_vline(xintercept = mean(coffee_data$daily_cups), color = "blue", size = 1) +
  labs(
    title = "Distribution of Daily Coffee Consumption",
    subtitle = "Red line: Hypothesized mean (3), Blue line: Sample mean",
    x = "Daily Cups of Coffee",
    y = "Count"
  ) +
  theme_minimal()

coffee_plot
```

```{r complete-infer-analysis-check}
grade_code("Outstanding! You've completed a comprehensive analysis using the infer workflow with clear interpretation and visualization.")
```

### Final Challenge: Complete infer Analysis

Now complete your own comprehensive analysis using the study hours data and the infer workflow:

```{r final-infer-challenge, exercise=TRUE}
# Your turn! Complete analysis of study hours data using infer
# Test H₀: μ = 20 vs H₁: μ ≠ 20

# Step 1: Summary statistics
study_summary <- study_hours %>%
  summarise(
    n = ___,
    mean_hours = ___,
    sd_hours = ___
  )

# Step 2: Calculate test statistic using infer
observed_t <- study_hours %>%
  specify(response = ___) %>%
  hypothesize(null = ___, mu = ___) %>%
  calculate(stat = ___)

# Step 3: Calculate p-value using infer
p_val <- study_hours %>%
  specify(response = ___) %>%
  hypothesize(null = ___, mu = ___) %>%
  calculate(stat = ___) %>%
  get_p_value(obs_stat = observed_t, direction = ___)

# Step 4: Make decision
alpha <- 0.05
decision <- ifelse(p_val$p_value < alpha, 
                   "Reject H₀", 
                   "Fail to reject H₀")

# Display all results
print("Summary Statistics:")
study_summary

print("Test Results:")
print(paste("T-statistic:", round(observed_t$stat, 4)))
print(paste("P-value:", round(p_val$p_value, 4)))
print(paste("Decision:", decision))
```

```{r final-infer-challenge-solution}
study_summary <- study_hours %>%
  summarise(
    n = n(),
    mean_hours = mean(hours_per_week),
    sd_hours = sd(hours_per_week)
  )

observed_t <- study_hours %>%
  specify(response = hours_per_week) %>%
  hypothesize(null = "point", mu = 20) %>%
  calculate(stat = "t")

p_val <- study_hours %>%
  specify(response = hours_per_week) %>%
  hypothesize(null = "point", mu = 20) %>%
  calculate(stat = "t") %>%
  get_p_value(obs_stat = observed_t, direction = "two-sided")

alpha <- 0.05
decision <- ifelse(p_val$p_value < alpha, 
                   "Reject H₀", 
                   "Fail to reject H₀")

print("Summary Statistics:")
study_summary

print("Test Results:")
print(paste("T-statistic:", round(observed_t$stat, 4)))
print(paste("P-value:", round(p_val$p_value, 4)))
print(paste("Decision:", decision))
```

```{r final-infer-challenge-check}
grade_code("Perfect! You've mastered the infer workflow for hypothesis testing. You can now confidently use this approach for statistical inference!")
```

## Summary

Congratulations! You've learned how to:

✅ **Use the infer grammar** for statistical inference  
✅ **Apply the specify-hypothesize-calculate workflow** for hypothesis testing  
✅ **Calculate test statistics and p-values** using a consistent framework  
✅ **Interpret results** and make statistical decisions  
✅ **Integrate infer with tidyverse tools** for complete analyses  
✅ **Create visualizations** to support your statistical conclusions  

### Key Takeaways

1. **The infer workflow is intuitive**: `specify() %>% hypothesize() %>% calculate()`
2. **Consistency across tests**: Same grammar works for different statistical tests
3. **Clear and explicit**: Each step makes your analysis transparent
4. **Tidyverse compatible**: Works seamlessly with dplyr and ggplot2
5. **Flexible framework**: Easy to extend to simulation-based methods

### The infer Advantage

The infer package makes statistical inference:
- **More intuitive** - the workflow mirrors statistical thinking
- **More consistent** - same grammar for different tests
- **More transparent** - each step is explicit
- **More flexible** - easy to switch between parametric and non-parametric approaches

### Next Steps

- Explore simulation-based approaches with infer
- Learn about confidence intervals using infer
- Apply infer to two-sample tests and categorical data
- Practice with different types of hypothesis tests using the same workflow

Great work completing this tutorial!